### 2021-01-17 papers/jin-thesis/Chapters/1Introduction.tex
Added a lot to my thesis directory

	Neuroegenerative diseases, such as Alzheimer's and Huntinton's disease, have consistently been one of the leading causes of death in the United States. Scientists all over the world struggle to find the leading causes of these diseases and continue to hope for a breakthrough in science in attempt to prevent or even cure these fatal diseases. Although this breakthrough has yet to come, researchers discovered that one of the central causes of cell degeneration is motor protein failure. More specifically, defects in the motor protein dynein has been linked to neurodegeneration due to the cell's funtional dependence on dynein and its mechanical properties. However, unlike dynein's motor protein siblings, dynein possesses a unique structure causing it to stochasitcally move around the cell and unpredicatable in nature. Thus, fully understandding the mechanisms of dynein can allow us to further conclude causes of its failure, putting us one step closer to solving the breakthrough.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\columnwidth]{Figures/dynein_walking_art.jpg}
	\caption[Artist Rendition of Dynein]{\textbf{Artist Rendition of Dynein} \cite{JohnsonArt}}
	\label{fig:final_disp}
\end{figure}

Cells are complicated. Despite their degeneracies, cells have the ability to move, perfectly divide, and spatially organize themselves. They are constructed by organelles surrdounded by a network of protein filaments called the cytoskeleton. The cytoskeleton is composed of microtubules (MT) that act as a walkway for motor protein soldiers to carry out the functions of the cell. These motor proteins consists of kinesin, myosin, and dynein.

@@

>>>>

	Neurodegenerative diseases, such as Alzheimer's and Huntinton's disease, have consistently been one of the leading causes of death in the United States. Scientists all over the world struggle to find the leading causes of these diseases and continue to hope for a breakthrough in science in attempt to prevent or even cure these fatal diseases. Although this breakthrough has yet to come, research suggests that motor protein failure is a main cause of cell degeneration. More specifically, defects in the motor protein dynein has been linked to neurodegeneration due to the cell's funtional dependence on dynein and its mechanical properties. However, unlike dynein's motor protein siblings, dynein possesses a unique structure causing it to stochasitcally move around the cell and be unpredicatable in nature. Thus, fully understanding the mechanisms of dynein can allow us to further conclude causes of its failure, putting us one step closer to discovering the major breakthrough for neurodegenerative research. In this paper, we propose a model of dynein that uses Monte Carlo methods and a Brownian dynamics simulation in order to reproduce the steping behavior observed from experiment. that reproduces the stepping behavior of experiment using Monte Carlo methods and a Brownian dynamics simulation.

@@

\begin{figure}[H]
	\centering
	\includegraphics[width=0.5\columnwidth]{Figures/dynein_walking_art.jpg}
	\caption[Artist Rendition of Dynein]{\textbf{Artist Rendition of Dynein} \cite{JohnsonArt}}
	\label{fig:ArtDynein}
\end{figure}

Cells are complicated. Despite their fatal degeneracies, cells have the ability to move, perfectly divide, and spatially organize themselves. They can do so because tiny, two-legged molecular proteins, named \textit{motor proteins}, carry cargo containing information that outlines the blueprint for different functions within the cell. These motor proteins consists of kinesin, myosin, and dynein. They can travel all around the cell on a cytoskeleton composed of microtubules (MT) that act as a walkway for the proteins. A depiction of dynein walking along the microtubule is shown above in Figure (\ref{fig:ArtDynein}). Microtubules are constructed from negatively charged $\alpha$ and positively charged $\beta$ tubulin dimers that polarizes the microtubule depending on which tubulin is left hanging at the ends. In a eukaryotic cell, microtubules typically stem outwards from the cell's nucleus to the cell walls where the $\beta$ tubulin is exposed, generating a positive end that pulls kinesin and myosin. However, dynein opposes kinesin and myosin by walking towards the nucleus cell body where the MT is negatively charged. This attraction to both ends drives the directed stepping motion for each motor protein and is shown in Figure (\ref{fig:transport}).

\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\columnwidth]{Figures/retrograde_transport.jpg}
	\caption[Retrograde Transport]{\textbf{Retrograde Transport} \cite{Vale2003molecular}}
	\label{fig:transport}
\end{figure}

Both kinesin and myosin exhibit hand-over-hand stepping, where their ``legs" take turns alternating steps like a human. Due to dynein's massive size compared to its motor protein siblings, dynein does not step uniformly but rather inchworms its ``feet" like a drunk human. This unique property makes dynein difficult to study when analyzing possible sources of molecular failure, but utters the importance of understanding the physical structure of the entire motor protein.

\section{Structure}
\textit{In-detail background knowledge on structure of dynein. Talk about each domain and their roles. Binding domain, six AAA+ domains, tail, and cargo. Stalk and linker that attaches things together. How the structure is so much larger than that of kinesin. dynein has flexibility because of its linker and in order to compensate for its huge domains. }

Dynein is composed of two motor heavy chain subunits of linked amino acids, called polypeptides, each which are separated into domains. These domains are the tail domain, the six AAA+ domains (which creates a ring and is often called the motor domain), and the microtubule binding domains, where a linker connects the tail to the AAA+ ring and a stalk connects the AAA+ ring to the binding domain. Figure (\ref{fig:structure}) displays a simplified dynein where both subunits are connected to the tail with separate linkers.

%\begin{figure}[H]
%	\centering
%	\includegraphics[width=0.6\columnwidth]{Figures/dynein_polypeptide.jpg}
%	\caption[Polypeptide Structure]{\textbf{Polypeptide Structure} \cite{GoodsellArt}}
%	\label{fig:polypeptide}
%\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=0.6\columnwidth]{Figures/dynein_on_MT.png}
	\caption[Dynein on Microtubule]{\textbf{Dynein on Microtubule} Artist’s illustration of a dimer of cytoplasmic dynein heavy chains bounded to microtubule. This rendition similarly illustrates dynein’s structure as a domain-rod system. \cite{TheTrappistArt}}
	\label{fig:structure}
\end{figure}

Dynein's crystal structure has been intensely studied because of the pivotal role it plays in dyneins unpredictable stepping nature. Its inability to step uniformly is largely due to the large lengths between each major domain site, as the linker and stalk are about a factor of a half larger than the diameter of the AAA+ ring. This is unique when commpared to kinesin and myosin, as their motor domains are much closer to the MT causing their center of mass to keep the protein controlled and steady when stepping. Despite dynein's large structure, dynein can still achieve processive stepping towards the negative end of the MT by the interaction of ATP with the AAA+ domains, thus is why the AAA+ domains are referred to as the motor domains.

\section{Stepping}
\textit{Biology and chemistry of stepping. What is actually causing it to step? ATP hydrolysis. Maybe short introduction on stepping and go more in detail in Power Stroke model. Say there are multiple theories and models of how dynein steps. Powerstroke, Winch, etc. Power stroke is most accepted. It can step on microtubule that is made up of tubulin dimers, $\alpha$ and $\beta$ tubulins. They are normally 8nm apart and so experimentalists believe dynein average step is always factor of 8nm. }

Dynein's ability to step is entirely ATP driven under the process of ATP hydrolysis with its motor domains.

\begin{figure}[H]
	\centering
	\includegraphics[width=0.3\columnwidth]{Figures/Onaxis.png}
	\caption[Stepping Axis]{\textbf{Stepping axis}  \cite{Dewitt2012} }
	\label{fig:YildizCorrelation}
\end{figure}

@@

\subsection{Powerstroke Model}
\textit{Theorized model of stepping. Mechanochemical Cycle. Talk about ATP hydrolysis. Different conformational changes within a step. Look at linker and how it changes angles. Physically, it looks like it unbinds, stretches leg, then kicks forward, then diffuse back to MT.}

\begin{figure}[H]
	\centering
	\includegraphics[width=1\columnwidth]{Figures/mechanochemical_cycle.jpeg}
	\caption[Mechanochemical Cycle]{\textbf{Mechanochemical Cycle}  \cite{Cianfrocco2015mechanism}}
	\label{fig:MechanochemicalCycle}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=1\columnwidth]{Figures/powerstroke.jpeg}
	\caption[Powerstroke]{\textbf{Powerstroke}  \cite{Carter2010communication} }
	\label{fig:Powerstroke}
\end{figure}

\section{Experimental Research on Dynein}
\textit{Most likely just short sections on experiments done on dynein.}
\subsection{Measuring}
\textit{How are experimentalists measuring dyneins stepping? Quantum dots on motor domains or quantum rod on tail, etc. }

\subsection{Interhead Coordination}
\textit{Purpose of my research. Trying to make model and fit to Yildiz analysis of dynein having interstep correlation. Dependency between step length and initial inter head separation.}

\begin{figure}[H]
	\centering
	\includegraphics[width=1\columnwidth]{Figures/Yildiz_stepping.png}
	\caption[Stepping Correlation]{\textbf{Stepping Correlation}  \cite{Dewitt2012} }
	\label{fig:YildizCorrelation}
\end{figure}
